Jug: Software for Reproducible Computation

### 2012-06-08 paper.tex
ENH Fix typos & improve text

In addition to the value to the community of being able to reproduce an
analysis, reproducible research practices allow for better control over the
project. For example, if necessary parameters to run a pipeline are kept
separately from the code that implements it (perhaps even in the researcher's
mind), this leads to error-prone analysis and opens up the possibility that
when the results are to be writtne up for publication, the researcher will no
longer be able to even completely describe the process that led to them.

For large projects, the use of multiple processors (either in the same machine
or distributed across a cluster) is necessary to obtain results in a useful
timeframe. Furthermore, it is often the case that, as the project evolves, it
becomes necessary to save intermediate results while down-stream analyses are
designed (or redesigned) and implemented. Therefore, having a single point of
entry (often in the form of a single script or a single main programme) for the
computation becomes increasingly difficult.

@@

Jug is written in pure Python, is completely cross-platform, and available as
free software under the MIT license. Jug is available from
\url{http://github.com/luispedro/jug}.
\bigskip
\bigskip

@@

These solutions do not necessarily scale to large problems, those that take
several days, months, or years of CPU time. For very large problems,
specialized solutions are needed, but there is a range of medium-sized problems
that can be successfully tackled on a computer cluster with a few hundred nodes
or less (so that a computation that would take a CPU-year can produce results
in less than a day).

@@

By allowing caching of results, where the caching depends explicitly on the
input parameters of that computation. Any change in the parameters immediately
triggers a recomputation of all dependendent results. The basic model is
similar to the one of the venerable Unix tool Make, which has been used before
for implementing reproducible research
pipelines~\citep{Schwab00makingscientific}.

However, unlike make, jug is written in Python and tasks can consist of any
Python function and its arguments. A jugfile (the file which specifies which
tasks need to be run---named by analogy to the makefiles of make) is a simple
Python script with some added notations (we will see just how little changes
need to be made from a conventional Python script).
\section{Design Choices}
\subsection{Task Based Architecture}

@@

Execution is the command used for running the tasks. In broad terms it performs
the loop below.

@@

The status command prints out a summary of the status of all of the tasks.
Figure~\ref{fig:jug-status-output} shows the output of this command using the
example jugfile above. We assume that the jugfile was called \code{images.py}
on disk and that there were 20~images in the directory. We can see that there
are 20~tasks ready to run, while the \code{mean} task is still waiting for the
results of the other tasks.
\subsection{Backends}

@@

Parallelisation is achieved by running more than one jug process
simmultaneously. All of the synchronisation is outsourced to the backend. This
has certain limitations, but it makes jug usable in certain cluster computing
environments with a batch architecture. As long as all of the processes can
access the backend, there is no need for them to communicate directly. It can
even be the case that processors start working on tasks in mid-processing.

This makes jug usable in batch-based computer cluster environments, which are
quite common in research institutions.

@@

It is often the case that one improves a step in a computational pipeline and
wished to obtain the result of recomputing the pipeline while reusing as much
of the previously computed intermediate results. With the manual management of
results I found this to be the most error-prone operation: by not removing
enough of the intermediate files, it was easy to generate an irreproducible
state where part of the computations were made using one version of the code
and part with another.

@@

It is still necessary to manually invalidate tasks. An alternative, which I
considered, was to have the code which implements the task be taken into
account while computing the hash. While this would add another layer of
protection, it would still be possible to make mistakes. If the function
depended on other functions, especially if this was done in a dynamic way, it
could be hard to discover all dependencies (cf.\ sumatra~\footnote{Sumatra is a
tool to collect all dependencies that are necessary to run a particular
computation. Its webpage is located at
\url{http://neuralensemble.org/trac/sumatra}.}). Additionally, minute
refactoring of the code would lead to over eager recomputation. This could make
the developer wary of making improvements in their code, resulting in overall
worse code.

@@

To evaluate classification, the dataset is broken up into 10~pieces and each of
the pieces is held-out of the analysis and then used to evaluate analysis (the
final result is the average of all ten values). This is known as
cross-validation. In this case, I use another library (which I also wrote)
which has a builtin method for jug-aware cross-validation.

@@

Jug has been available in the Python Package Index\footnote{The Python Package
Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
through February~2012 and has been downloaded over 10,000~times (some of these
downloads may, of course, represent upgrades, and users which downloaded it
from some other source will not be counted).

>>>>

In addition to being valuable to the wider community, which may wish to
reproduce or extend a published analysis, reproducible research practices allow
for better control over the project by the authors themselves. For example, if
necessary parameters to run a pipeline are kept separately from the code that
implements it (perhaps even in mind of the researcher), this leads to
error-prone analysis and opens up the possibility that when the results are to
be written up for publication, the researcher will no longer be able to even
completely describe the process that led to them.

For large projects, the use of multiple processors (either in the same machine
or distributed across a cluster) is necessary to obtain results in a useful
time frame. Furthermore, it is often the case that, as the project evolves, it
becomes necessary to save intermediate results while down-stream analyses are
designed (or redesigned) and implemented. Under many frameworks, this causes
having a single point of entry (often in the form of a single script or a
single main programme) for the computation becomes increasingly difficult.

@@

Jug is written in pure Python, is completely cross-platform, and available as
free software under the liberal MIT license. Jug is available from
\url{http://github.com/luispedro/jug}.
\bigskip
\bigskip

@@

These solutions do not necessarily scale to large problems, those that take
several days, months, or years of CPU time. For very large problems,
specialized solutions are needed~\citep{mapReduce}. There is a range of
medium-sized problems that can be successfully tackled on a computer cluster
with a few hundred nodes or less (so that a computation that would take a
CPU-year can produce results in less than a day).

@@

By allowing caching of results, where the caching depends explicitly on the
input parameters of that computation. Any change in the parameters immediately
triggers a recomputation of all dependent results. The basic model is similar
to the one of the venerable Unix tool Make, which has been used before for
implementing reproducible research pipelines~\citep{Schwab00makingscientific}.

However, unlike make, jug is written in Python and tasks can consist of any
Python function and its arguments. This can include running external commands.
A jugfile (the file which specifies which tasks need to be run---named by
analogy to the makefiles of make) is a simple Python script with some added
notations (we will see just how little changes need to be made from a
conventional Python script).

\section{Design Choices}
\subsection{Task Based Architecture}

@@

Execution is the command used for running the tasks. It performs a more complex
version of this loop:

@@

The status command prints out a summary of the status of all of the tasks.
Figure~\ref{fig:jug-status-output} shows the output of this command using the
example jugfile above. We assume that the jugfile was called \code{images.py}
on disk and that there were 20~images in the directory. We can see that there
are 20~tasks ready to run, while the \code{mean} task is still waiting for the
results of the other tasks.

\subsection{Backends}

@@

Parallelisation is achieved by running more than one jug process
simultaneously. All of the synchronisation is outsourced to the backend. As
long as all of the processes can access the backend, there is no need for them
to communicate directly. It can even be the case that processors start working
on tasks in mid-processing. This makes jug usable in batch-based computer
cluster environments, which are quite common in research institutions.

@@

It is often the case that one improves a step in a computational pipeline and
wished to obtain the result of recomputing the pipeline while reusing as much
of the previously computed intermediate results. Before using Jug, I found this
to be a very error-prone operation: by not removing enough of the intermediate
files, it was easy to generate an irreproducible state where part of the
computations were made using one version of the code and part with another.

@@

It is still necessary to manually invalidate tasks. An alternative, which I
considered, was to have the code which implements the task be taken into
account while computing the hash. While this would add another layer of
protection, it would still be possible to make mistakes. If the function
depended on other functions, especially if this was done in a dynamic way, it
could be hard to discover all dependencies (cf.\ sumatra~\footnote{Sumatra is a
tool to collect all dependencies that are necessary to run a particular
computation. Its webpage is located at
\url{http://neuralensemble.org/trac/sumatra}.}). Additionally, even minute
refactoring of the code would lead to over eager recomputation. This could make
the developer wary of making improvements in their code, resulting in overall
worse code.

@@

To evaluate classification, the dataset is broken up into 10~pieces and each of
the pieces is held-out of the analysis and then used to evaluate analysis (the
final result is the average of all ten values). This is known as
cross-validation. In this case, I use another library (which I also wrote)
which has a built-in method for jug-aware cross-validation.

@@

Jug has been available in the Python Package Index\footnote{The Python Package
Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
through June~2012 and has been downloaded over 13,000~times (some of these
downloads may, of course, represent upgrades, and users which downloaded it
from some other source will not be counted).
